### 2015-08-31 Projects/DelCo_2014/Report/Hose_Stream_Report/Report/DelCo_Hose_Tests_2014.tex
DelCo_2014: Updates to hose stream report and plotting/sorting scripts

This report is concerned with examining the impact of hose stream selection and the pattern in which the stream is applied on the ventilation inside a structure with various flow path configurations. Two types of experiments were conducted in order to examine the impact. First, cold flow tests were performed. These tests involved using a 0.46 m (18 in) diameter positive pressure ventilation (PPV) fan to move air through a structure with various flow path configurations. The second type of experiment involved flowing water using various stream and application patterns into a structure with flow path configurations similar to the cold flow test configurations. Air velocities measured during the hose flow experiments were compared to the velocities measured during the cold flow tests to determine the impact of hose stream and application pattern combinations on ventilation throughout the structure relative to the impact of using a 0.46 m (18 in) diameter PPV fan to move air throughout the structure.

@@

\section{Background}
\label{sec:Background}
%http://www.tft.com/literature/library/files/bfa.pdf
Bill Nelson and Keith Royer strongly recommend that the rotation be clockwise as viewed from the nozzle operator’s position. From many experiments it was determined that a clockwise rotation is more effective than a counterclockwise rotation. They say that:
1. Clockwise rotation is safer since it drive smoke, gases, and flames away from the nozzle.
Counterclockwise rotation does just the opposite.
2. Clockwise rotation produces steam with an active rolling action. Counterclockwise rotation
produces steam with an inactive and lazy action.
3. A clockwise rotation produces a faster knockdown time.

@@

\subsection{Construction}
\label{sec:Construction}
Two concrete structures were built on a concrete slab as shown in Fig.~\ref{fig:struct_pics}. They were designed to simulate a single-story and a two-story residential structure.  The first floor of each structure had an outer wall composed of interlocking concrete blocks of dimensions 0.61 m x 0.61 m 0.61 m (2 ft x 2ft x 2ft). The interior dimensions of each structure were 6.1 m (20 ft) wide, 11 m (36 ft) long and 2.4 m (8 ft) high.  The joints and gaps between the blocks were filled with high temperature insulation.

The interior walls of the first floor of each structure were framed with steel studs and track.  The studs were set to 0.40 m (16 in) centers.  The ceiling/floor support was composed of wood truss joist I-beams (TJIs) with a 299 mm (11.75 in) depth.  The TJI was composed of laminated veneer lumber flanges with a cross section of 29 mm (1.125 in) x 44 mm (1.75 in) and an 11 mm (0.43 in) thick oriented strand board web as shown in Figure  XX .  Tongue and grove, 18.3 mm (0.72 in) thick, oriented strand board was screwed (nailed?) to the top of the TJIs.

The second story of the two level structure was built on the wood floor assembly as described above. The walls were framed with wood studs etc...

The interior walls of the burn room were lined with 13 mm (0.5 in) thick cement board.  The ceiling of the burn room was the exposed ``floor assembly".  The walls of the hallway and entry foyer were composed of 16 mm (0.625 in) Type X gypsum room. The ceiling of the hallway and entry foyer was composed of two layers of 13 mm (0.5 in) thick cement board.

@@

\subsection{Instrumentation}
\label{sec:Instrumentation}
Schematic plan overviews of the instrumentation in the East and West Structures are provided in Fig.~\ref{fig:east_instrumentation} and Fig.~\ref{fig:west_instrumentation}. There is a discussion of uncertainties for each measurement below in Section~\ref{sec:Uncertainty}. Gas velocity was measured at various doorways using differential pressure transducers connected to bi-directional velocity probes~\cite{McCaffrey:Combustion_and_Flame}. Each set of BDPs contained 8 probes located at distances of 0.08 m, 0.34 m, 0.61 m, 0.88 m, 1.15 m, 1.42 m, 1.68 m, and 1.95 m below the soffit of the corresponding doorway. Fig.~\ref{fig:BDPs} shows a picture of two sets of BDPs, A5 and A6, located at each double door on the first floor of the West Structure. A single thermocouple was attached to each bi-directional probe. The thermocouples that are used with the bi-directional probes are exposed-bead, Chromel-Alumel (type K) with a 1.0 mm (0.04 in) diameter. Starting with the exposed bead, the thermocouple wire was sheathed in a 3.2 mm (0.13 in) diameter Inconel shield, 0.76 m (2.5 ft) in length.

@@

\begin{figure}[!ht]
\includegraphics[width=6in]{../Pictures/BDPs}
\caption[Two sets of BDPs in Doorway of West Structure]{Two sets of BDPs, A5 and A6, in each doorway of the double doors on the first floor of the West Structure}
\label{fig:BDPs}
\end{figure}

@@

Each length measurement was taken carefully. Length measurements such as the room dimensions,
instrumentation array locations, and fire apparatus (for example nozzle, sprinkler, or fan) placement were made with a hand held laser measurement device which is has an accuracy of $\pm6.0$ mm (0.24 in) over a range of 0.61 m (2.00 ft) to 15.3 m (50.0 ft)~\cite{Stanley}. However, conditions affecting the measurement, such as levelness of the device, yield an estimated uncertainty of $\pm$0.5\% for measurements in the 2.0 m (6.6 ft) to 10.0 m (32.8 ft) range. Steel measuring tapes with a resolution of $\pm$0.5 mm were used to locate individual sensors within a measurement array and to measure and position the furniture. The steel measuring tapes were manufactured in compliance with NIST Manual 44, which specifies a tolerance of $\pm1.6$ mm (0.06 in) for 9.1 m (30 ft) tapes and $\pm6.4$ mm (0.25 in) for 30.5 m (100 ft) tapes \cite{NIST_Manual_44}. Some issues, such as "soft" edges on the upholstered furniture, result in an estimated total expanded uncertainty of $\pm$1.0\%.

Bi-directional probes and single thermocouples were used to measure the velocity. The bi-directional probes used similar pressure transducers as those used for the differential pressure measurements discussed above. Bare-bead Type K thermocouple are co-located with the probe. A gas velocity measurement study, examining the doorway flow of pre-flashover compartment fires, yielded expanded uncertainty measurements ranging from $\pm0.14$ to $\pm0.22$ for bi-directional probes of similar design \cite{Bryant:FSJ2009}. The total expanded uncertainty for gas velocity in these experiments estimated to be $\pm18$\%.

Water flowrate was measured with a pressure and flow meter combination shown in Fig. \ref{fig:flow_meter}. The meter consists of a section of 6.35 cm (2.5 inch) cast aluminum pipe with a 0-4.1 MPa (0-600 psi) pressure transducer and a paddlewheel type flow sensor with a range of 0 to 4800 lpm (1250 gpm). The pressure transducer and paddlewheel both connect to the battery operated control box where the pressure transducer voltage is converted to a pressure and the paddlewheel pulse count is converted to a volumetric flow rate. The manufacturer reports a $\pm5$\% calibration expanded uncertainty for the flow sensor and $\pm3$\%  for the pressure sensor \cite{Akron}. The pressure transducer was calibrated with a known analog pressure gauge. The flow meter was calibrated by capturing water over time and measuring that mass of water to determine the flowrate. The total expanded uncertainty was estimated at $\pm10$\%.

@@

\subsection{Cold Flow Experimental Procedure}
\label{sec:Cold_Flow_Procedure}

\subsection{Water Flow Experimental Procedure}
\label{sec:Water_Flow_Procedure}
A variety of experiments were conducted to examine the impact of different hose streams and application methods on air flow inside a structure. All the experiments were conducted in either the East or West Structure; used a monitor %reference for specs? 2000 LPM (500 GPM)?
or 1 \nicefrac{3}{4}" handline with a combination nozzle (Fig.~\ref{fig:monitor+handline}) to flow water at approximately 120 GPM in a straight, narrow fog, or wide fog stream pattern; and involved opening and closing doors in the structure to change the ventilation patterns in the structure. Experiments that used a monitor to flow water focused on varying the location of water application, while the experiments that used a handline to flow water focused on varying the pattern in which water was applied.

@@

\section{Cold Flow Test Results}
\label{sec:Cold_Flow_Test_Results}

@@

>>>>

This report is concerned with examining the impact of hose stream and application pattern selection on the ventilation inside a structure with various flow path configurations. Experiments in which water was flowed using various stream and application patterns into a structure with different flow path configurations were conducted. Air velocities were measured and used to determine the impact of hose stream and application pattern combinations on ventilation throughout the structure.

@@

\section{Background}
\label{sec:Background}
%http://www.tft.com/literature/library/files/bfa.pdf
Nelson and Royer suggest using a clockwise rotation instead of a counterclockwise application pattern for the following reasons:
\begin{enumerate} 
\item Clockwise rotation is safer since it drive smoke, gases, and flames away from the nozzle.
Counterclockwise rotation does just the opposite.
\item Clockwise rotation produces steam with an active rolling action. Counterclockwise rotation
produces steam with an inactive and lazy action.
\item A clockwise rotation produces a faster knockdown time.
\end{enumerate}

@@

\subsection{Construction}
\label{sec:Construction}
Two concrete structures were built on a concrete slab as shown in Fig.~\ref{fig:struct_pics}. They were designed to simulate a single-story and a two-story residential structure. The first floor of each structure had an outer wall composed of interlocking concrete blocks of dimensions 0.61 m x 0.61 m X 0.61 m (2 ft x 2 ft x 2 ft). The interior dimensions of each structure were 6.1 m (20 ft) wide, 11 m (36 ft) long and 2.4 m (8 ft) high. The joints and gaps between the blocks were filled with high temperature insulation.

The interior walls of the first floor of each structure were framed with steel studs and track. The studs were set to 0.40 m (16 in) centers. The ceiling/floor support was composed of wood truss joist I-beams (TJIs) with a 299 mm (11.75 in) depth. The TJI was composed of laminated veneer lumber flanges with a cross section of 29 mm (1\nicefrac{1}{8} in) x 44 mm (1.75 in) and an 11 mm (0.43 in) thick oriented strand board web as shown in Figure \ref{fig:TJI}. Tongue and grove, 18.3 mm (0.72 in) thick, oriented strand board was screwed to the top of the TJIs.

\begin{figure}[!ht]
\includegraphics[width=6in]{../Pictures/TJI_support}
\caption[Truss Joist I-Beams]{Truss joist I-beams used for the ceiling/floor support in the West Structure}
\label{fig:TJI}
\end{figure}

The second story of the two level structure was built on the wood floor assembly as described above. The walls were framed with wood studs..**** NEED MORE DETAILS ******

The interior walls of the burn room were lined with 13 mm (0.5 in) thick cement board.  The ceiling of the burn room was the exposed "floor assembly".  The walls of the hallway and entry foyer were composed of 16 mm (\nicefrac{3}{8} in) Type X gypsum room. The ceiling of the hallway and entry foyer was composed of two layers of 13 mm (0.5 in) thick cement board.

@@

\subsection{Instrumentation}
\label{sec:Instrumentation}
Schematic plan overviews of the instrumentation in the East and West Structures are provided in Fig.~\ref{fig:east_instrumentation} and Fig.~\ref{fig:west_instrumentation}. There is a discussion of uncertainties for each measurement below in Section~\ref{sec:Uncertainty}. Gas velocity was measured at various doorways using differential pressure transducers connected to bi-directional velocity probes (BDPs)~\cite{McCaffrey:Combustion_and_Flame}. Each set of BDPs contained 8 probes located at distances of 0.08 m, 0.34 m, 0.61 m, 0.88 m, 1.15 m, 1.42 m, 1.68 m, and 1.95 m below the soffit of the corresponding doorway. Fig.~\ref{fig:BDPs} shows a picture of two sets of BDPs, A5 and A6, located at each double door on the first floor of the West Structure. A single thermocouple was attached to each bi-directional probe. The thermocouples that are used with the bi-directional probes are exposed-bead, Chromel-Alumel (type K) with a 1.0 mm (0.04 in) diameter. Starting with the exposed bead, the thermocouple wire was sheathed in a 3.2 mm (0.13 in) diameter Inconel shield, 0.76 m (2.5 ft) in length.

@@

\begin{figure}[!ht]
\includegraphics[width=6in]{../Pictures/BDPs}
\caption[Sets of Bi-directional Velocity Probes in Doorway of West Structure]{Two sets (A5 and A6) of bi-directional velocity probes in each doorway of the double doors on the first floor of the West Structure}
\label{fig:BDPs}
\end{figure}

@@

Each length measurement was taken carefully. Length measurements such as the room dimensions,
instrumentation array locations, and fire apparatus (for example nozzle, sprinkler, or fan) placement were made with a hand held laser measurement device which is has an accuracy of $\pm6.0$ mm (0.24 in) over a range of 0.61 m (2.00 ft) to 15.3 m (50.0 ft)~\cite{Stanley}. However, conditions affecting the measurement, such as levelness of the device, yield an estimated uncertainty of $\pm$0.5\% for measurements in the 2.0 m (6.6 ft) to 10.0 m (32.8 ft) range. Steel measuring tapes with a resolution of $\pm$0.5 mm were used to locate individual sensors within a measurement array and to measure and position the furniture. The steel measuring tapes were manufactured in compliance with NIST Manual 44, which specifies a tolerance of $\pm1.6$ mm (0.06 in) for 9.1 m (30 ft) tapes and $\pm6.4$ mm (0.25 in) for 30.5 m (100 ft) tapes~\cite{NIST_Manual_44}. Some issues, such as "soft" edges on the upholstered furniture, result in an estimated total expanded uncertainty of $\pm$1.0\%.

Bi-directional probes and single thermocouples were used to measure the velocity. The bi-directional probes used similar pressure transducers as those used for the differential pressure measurements discussed above. Bare-bead Type K thermocouple are co-located with the probe. A gas velocity measurement study, examining the doorway flow of pre-flashover compartment fires, yielded expanded uncertainty measurements ranging from $\pm0.14$ to $\pm0.22$ for bi-directional probes of similar design~\cite{Bryant:FSJ2009}. The total expanded uncertainty for gas velocity in these experiments estimated to be $\pm18$\%.

Water flowrate was measured with a pressure and flow meter combination shown in Fig.~\ref{fig:flow_meter}. The meter consists of a section of 6.35 cm (2.5 inch) cast aluminum pipe with a 0-4.1 MPa (0-600 psi) pressure transducer and a paddlewheel type flow sensor with a range of 0 to 4800 lpm (1250 gpm). The pressure transducer and paddlewheel both connect to the battery operated control box where the pressure transducer voltage is converted to a pressure and the paddlewheel pulse count is converted to a volumetric flow rate. The manufacturer reports a $\pm5$\% calibration expanded uncertainty for the flow sensor and $\pm3$\%  for the pressure sensor~\cite{Akron}. The pressure transducer was calibrated with a known analog pressure gauge. The flow meter was calibrated by capturing water over time and measuring that mass of water to determine the flowrate. The total expanded uncertainty was estimated at $\pm10$\%.

@@

% \subsection{Cold Flow Experimental Procedure}
% \label{sec:Cold_Flow_Procedure}

% \subsection{Water Flow Experimental Procedure}
% \label{sec:Water_Flow_Procedure}
A variety of experiments were conducted to examine the impact of different hose streams and application methods on air flow inside a structure. All the experiments were conducted in either the East or West Structure; used a monitor %reference for specs? 2000 LPM (500 GPM)?
or 1~\nicefrac{3}{4}" handline with a combination nozzle (Fig.~\ref{fig:monitor+handline}) to flow water at approximately 120 GPM in a straight, narrow fog, or wide fog stream pattern; and involved opening and closing doors in the structure to change the ventilation patterns. Experiments that used a monitor to flow water focused on varying the location of water application, while the experiments that used a handline to flow water focused on varying the pattern in which water was applied.

@@

@@

\begin{table}[!ht]
%\addtolength{\tabcolsep}{-3.25pt}
%\small
\caption{Average air velocity (m/s) through Door A6 for tested combinations of stream pattern and application location during clockwise and counter clockwise rotations}
\begin{tabular}{lcccc}
\toprule
 & \multicolumn{2}{c}{\underline{Room A}} & \multicolumn{2}{c}{\underline{Room B Ceiling}}
\\
\textbf{Stream} & \textbf{Clockwise} & \begin{tabular}{@{}c@{}} \textbf{Counter} \\ \textbf{Clockwise} \\ \end{tabular} & \textbf{Clockwise} & \begin{tabular}{@{}c@{}} \textbf{Counter} \\ \textbf{Clockwise} \\ \end{tabular}
\\
\midrule 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textit{Straight} 	& 
%%%% Room A %%%% 
0.6 $\pm 0.3$ & 0.8 $\pm 0.3$ 
%%%% Room B Ceiling %%%% 
& 0.5 $\pm 0.3$ & 0.5 $\pm 0.3$
\\	\multicolumn{5}{c}{}	\\
\textit{Narrow Fog} & 
%%%% Room A %%%% 
1.8 $\pm 0.3$ & 1.6 $\pm 0.3$
%%%% Room B Ceiling %%%% 
 & 1.5 $\pm 0.3$ & 1.6 $\pm 0.3$
\\	\multicolumn{5}{c}{}	\\
\textit{Wide Fog} 	& 
%%%% Room A %%%% 
 2.1 $\pm 0.3$ & 2.2 $\pm 0.3$
%%%% Room B Ceiling %%%% 
 & 2.3 $\pm 0.3$ & 2.6 $\pm 0.3$ \\
\bottomrule
\end{tabular}
\label{table:Test_35_CW_CCW}
\end{table}
